Privacy-Respecting Type Error Telemetry at Scale

### 2024-01-31 programming23/paper.tex
proofreading

{Roblox} is a platform for {shared virtual experiences} (typically 3D games), with
70~million Daily Active Users, and 16~billion hours of engagement in
July--September 2023~\cite{corp.roblox.com}.
There are over 5~million distinct user-created programs available on the platform thanks to a
worldwide community of 3~million creators.
Many of the creators program for fun or learning and may
not consider themselves software developers; others are professional
developers.

@@

Due to the broad community of creators, the goals of the Luau type
system are rather unique~\cite{bfj-hatra-2021}.
Whereas a traditional type system focuses on compilation and memory
safety, Luau takes a lenient approach by default and lets creators
gradually~\cite{st-sfp-2006,tf-popl-2008} migrate to rigorous checks
one module at a time.
The \mnocheck{} mode reports only syntax errors,
the \mnonstrict{} mode reports would-be runtime errors and skips over
other issues,
and the \mstrict{} mode reports a variety of potential errors.
In addition, no matter the mode, a {\FS{}} type analysis runs to guide \IDE{} tools~(\cref{s:context}).
Luau's types aim to support untyped designs, in the spirit of migratory
typing~\cite{tfffgksst-snapl-2017}, so that creators can switch modes without
needing to restructure their code in major, potentially-breaking, ways.

@@

In this work, we design telemetry that collects data on type errors without
exposing source code, source locations, or even error message text (which
may contain revealing information).
The telemetry counts the number and \emph{kind} of type errors at various
levels of granularity.
Furthermore it maintains a client-side approximation of the latest source-code
edits and uses that to identify type errors that overlap with this edit range.
Telemetry records are correlated by session, using a \panon{} session
identifier and a timestamp.

With this telemetry data, we investigate research questions about
the adoption and benefits of type analysis:
\begin{description}
  \item[RQ1.]
    How many sessions use type analysis?
    How often do projects contain modules with different
    analysis modes?
    How often do sessions turn analysis off?
  \item[RQ2.]
    For modules that use type analysis:
    which errors arise,
    how do sessions respond,
    and which errors tend to persist despite subsequent edits?
  \item[RQ3.]
    What impact does type analysis have on the number of \FS{}
    errors?
    For example, do \FS{} errors pile up in unanalyzed projects?
\end{description}

@@

At a higher level, this paper is the first to use telemetry 
to study a type checker.
It thus represents a step toward data-driven language design,
informed by many users' actual practice.
Our data captures over 340,000 sessions
that occured between February and April 2023
and covers thousands of type analysis errors
and millions of \FS{} errors.
By contrast to typical qualitative methods such as surveys and interviews, it
is not restricted to users' \emph{perceptions} about their work and it is not
limited to a small number of users.

@@

  \item
    An available dataset of over 1.5 million telemetry records
    and scripts to analyze them: \url{https://zenodo.org/doi/10.5281/zenodo.10275213}

@@

Creators can opt in to detailed error reports and highlights by
choosing a \emph{type analysis mode} for each script they write:
\begin{itemize}
  \item \mnocheck{}: report only syntax errors (the default),
  \item \mnonstrict{}: report syntax errors and a subset of high confidence type errors, or
  \item \mstrict{}: report syntax errors and all type errors.
\end{itemize}
Each run of type analysis can report several errors.
There is no guarantee that creators read every error.
In fact, creators who close the Script Analysis widget
will see highlights in their code but no further details.

@@

 As an example of \mnonstrict{} mode, the following program only reports one error:
\begin{verbatim}
  --!nonstrict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> OK
  local z = x.r           --> UnknownProperty: Key 'r' not found in table 'x'
\end{verbatim}
In \mstrict{} mode, it reports two errors:
\begin{verbatim}
  --!strict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> TypeMismatch: Type 'nil' could not be converted into 'number'
  local z = x.r           --> UnknownProperty: Key 'r' not found in table 'x'
\end{verbatim}
In cases like this, where it is undecidable whether there will be a run-time error,
\mstrict{} mode errs on the side of reporting an error, and \mnonstrict{} mode errs on
the side of suppressing the error.

Both modes report the \code{UnknownProperty} error because
misspellings of property names are common enough to report in both
\mstrict{} and \mnonstrict{} mode.
\mnocheck{} reports no error because the program is syntactically valid.
See~\cite{bfj-hatra-2021} for a more detailed discussion of the rationale
behind the type systems.

In addition to the main type analysis, Roblox Studio takes a second pass over
every codebase with a \emph{\FS{}} analysis to infer autocomplete suggestions
and drive other \IDE{} tools.
To infer precise types,
the \FS{} analysis uses rigorous checks very similar to those of
\mstrict{} mode (but not identical: see \cref{s:strict-vs-forcedstrict})
no matter what mode the creator declared for the script.
(Internally to Roblox Studio, this analysis is called \emph{forced strict}
because it ignores the creator's choice and applies strong checks.)
When it fails to infer a type, \FS{} analysis defaults to the unknown gradual
type~\cite{st-sfp-2006}.
Creators cannot see the errors reported by \FS{} analysis.

\Cref{t:type-error-labels} lists a few of the type errors that
script analysis can report (10 out of 35 total)
to give a sense of Luau.
A \code{SyntaxError} is the only error that can appear in
\mnocheck{} code.
Several errors in addition to \code{UnknownProperty} are about
tables, which are Luau's (and Lua's) primary data structure.
Tables support arrays, dictionaries, and objects; consequently,
tables can have methods and can usually be extended.
The error \code{TypesAreUnrelated} is a special kind of type mismatch
that arises only during cast, unification, or subtyping check.
A \code{CountMismatch} occurs when the arguments to a function
do not match the function's arity.
Similarly, an \code{IncorrectGenericParamCount} is an arity mismatch
for a generic type.
Lastly, \code{CodeTooComplex} occurs when the typechecker hits an
internal limit on problem size and \code{GenericError} is a catch-all
label for miscellaneous errors.
When Luau can provide more context regarding a generic error, it uses
a second error label called \code{ExtraInformation} to attach
a brief description.

@@

  \begin{tabular}{ll}
    Label & Interpretation \\\midrule
    \code{TypeMismatch} & Basic type error. \\
    \code{SyntaxError} & Basic parse error, e.g., \code{for if end}. \\
    \code{UnknownProperty} & Referenced an invalid field or method.  \\
    \code{OnlyTablesCanHaveMethods} & Tried to attach a method to a non-table. \\
    \code{CannotExtendTable} & Tried to extend a sealed table. \\
    \code{TypesAreUnrelated} & Failed cast, unify, or subtype. \\
    \code{CountMismatch} & Arity mismatch for a function. \\
    %% \code{CannotCallNonFunction} & Called a value that is not a function. \\
    \code{IncorrectGenericParamCount} & Arity mismatch for a generic type. \\
    \code{CodeTooComplex} & Type analysis failed~(\cref{s:code-too-complex}). \\
    %% \code{UnificationTooComplex} & Type analysis failed, unification solver hit a limit \\
    \code{GenericError}
    & Generic label for other non-type errors,\\
    & \hbox{}~~such as looping over an unordered table.
    %% attempting to extend a type that does not describe a class
    %% attempting to iterate over a table without a clear order

@@

Luau modules provide reusable libraries, which may be
{imported} by other scripts.
Imports therefore form a graph in general, though the analysis
removes edges to ensure an acyclic graph.
In \mstrict{} mode, it is an error to create an import cycle.

@@

While a program is under development, it is typical for the data
model to be edited.
% For example, instances may be added or renamed.
% --> Commented out to get a better page break.
Since the shape of the data model tree is reflected in
the type system, it is possible for these edits to introduce a mountain
of type errors across the project.

@@

Telemetry allows an application to ``phone home'' with data
summarizing usage patterns, such performance data, crash reporting,
and feature uptake. A typical usage is in deciding whether an API can
be deprecated; without telemetry, it may be difficult to know how
popular the API is, but with telemetry it is straightforward.
Telemetry for programming languages can, however, be controversial.
See, for example, the lively discussion around telemetry in the Go
toolchain~\cite{golang-telemetry}.

@@

\end{enumerate}
See Cox~\cite{transparent-telemetry} for further discussion of
the privacy implications and tradeoffs in telemetry.

@@

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    \uid{}: a \panon{} identifier (15-digit random number) for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summary of type errors and \FS{} errors during the
    last two invocations of type analysis, and the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: list of specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

@@

\Cref{t:session-size} similarly outlines the size of sessions using two metrics:
real time in seconds between the first and last record, and the total number of
records.
The plots focus
on the lower-left fragment of the data, which again has a very long tail.
%% comment on long-tail record counts?
The longest sessions last several days and/or contain thousands of records,
while the shortest last a few milliseconds or include just one record.
%% 0 ms = 1 event
A median session runs on the order of minutes and consists of a few dozen events.

@@

The low fraction of errors in the edit range sends a mixed message:
\begin{itemize}
  \item
    On one hand, it may be that edits successfully remove errors.
    This is true, however, only for type errors and not for \FS{} errors.
    There are 11,479 more type errors from the previous analysis
    that overlap with the current edit range~(\pct{27} difference),
    but 172,678 fewer previous \FS{} errors~(\pct{15} difference).
    Since creators can see the type errors but not the \FS{}
    errors before making edits, we conclude that the visibility
    makes a difference.
    \Cref{s:type-error-survival} explores edits and type errors in more detail.

  \item
    On the other hand, the low overlap between errors (whether current or previous)
    and edits says that edits rarely target highlighted code.
    There are several possible explanations:
    analysis might report several errors and the edits only target one or two of them;
    creators might ignore errors entirely and edit other code; or
    creators might choose to edit a non-highlighted location to fix an error.
    While non-privacy-protecting analyses have answered such
    questions~\cite{mfk-sigcse-2011}, our data cannot narrow down the answer.

@@

%% mostly pure, rare mixes
Most sessions (\pct{99.82}) stick to a single analysis mode.
They never change the mode in the current module and never
switch focus to a module with a different analysis mode.
Among the other sessions, most of those (\pct{80}) switch to a
module with a different mode, about half contain at least one edit that
upgrades to a more strict mode, and about half contain a downgrade to a
less-strict mode.
There are 263 total upgrades (across all sessions) and 320 total downgrades.

@@

The data does not support this explanation.
On average, mode-upgrades resulted in 3 additional
type errors (stddev 7, median 0).
This data excludes module switch records and considers only keystroke events.
The worst-case increase was quite high at 57 type errors, but exceptional.
In another exception, upgrading modes removed three type errors,
possibly due to other edits being rolled in with the
mode change.

@@

\Cref{f:error-by-mode} divides errors across modes.
By the left plot, \mnonstrict{} accounts for two thirds of all type errors.
This confirms our expectation.
%% because \mnocheck{} reports only syntax errors and
%% \mstrict{} covers only a small number of telemetry records.
The right plot, by contrast, does not support our hypothesis that \mnocheck{}
records have more \FS{} errors than the other modes.
Though \mnocheck{} does have the most \FS{} errors, its \emph{proportion} matches
the proportion of \mnocheck{} records in the
dataset~(\cref{f:total-records})---if every record had exactly one \FS{} error,
the plot would be the same.

@@

The main focus of our telemetry design is to learn how type errors intersect
with edit ranges.
Indeed, most of the fields in a telemetry record are dedicated to this
topic~(\cref{f:tdata}), and are directed to RQ2 on type errors .

@@

    Curiously, the number of syntax errors never decreases or stays the same.
    This may be because the easy fix is to press undo, and telemetry cannot
    see undo events.

    %% TODO awkward
    Our telemetry strategy likely inflates the number of typos because it cannot
    determine when a creator is mid-edit.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will generate an
    \code{UnknownSymbol} error for \code{buck} and
    \code{UnknownProperty} for \code{bucket.c}, which may end up getting sampled.
    \Cref{s:conclusion} suggests ways
    to avoid this in the future.

  \item
    After the syntax errors, \code{CountMismatch}, \code{TypeMismatch}, and
    \code{GenericError} are the next most common \mnonstrict{} and \mstrict{}
    errors.
    A \code{CountMismatch} sends the wrong number of arguments to a function.
    It may simply be a common error, but it may also be a
    telemetry artifact.
    The high incidence of \code{TypeMismatch} and \code{GenericError} calls for
    further study in future work to determine precisely which error arose.
    (Some of the \code{GenericError}s come with context strings in the form of
    \code{ExtraInformation} errors, but telemetry does not report these strings.)
    %% GenericError = only a few things: extend non-class, iterate non-table, ...?

@@

  \item
    \code{TypeMismatch} is far more common in \mstrict{} (\pct{20.46})
    than \mnonstrict{} (\pct{2.44}) because \mstrict{} mode
    has additional type constraints.
    Several other errors have higher \mstrict{} percentages for the
    same reason, such as \code{CannotInferBinaryOperation}, and
    two errors appeared only in \mstrict{}:
    \code{OccursCheckFailed} and \code{IncorrectGenericParamCount}.

  \item
    Several Luau errors never appeared in edit ranges.
    Four of these deal with internal typechecker limits, which we discuss
    below.
    The others are for: dynamic property lookups,
    deprecated APIs, swapped generic parameters,
    functions that require a self, and type pack mismatches.
    We would not expect these to appear often.

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{ktu-caap-1990,m-popl-1990}, the typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors along with a message
that the analysis failed to understand the code:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
These errors can appear in Script Analysis just like any other
error, but one design goal of Luau is that code rarely hits the limits.
A typical creator should never see these errors.

@@

Since errors and edits may not overlap, either because the highlights
are misleading or because the creator ignores the errors, it is important
to see how the total errors change over time.
We measure \emph{error density} rather than error count, however,
to normalize for the wide range of codebase sizes~(\cref{t:codebase-size}).
Density is the number of type errors divided by the number of lines.

\Cref{f:error-density} plots changes to error density for every session in the dataset.
The data excludes module-switch telemetry records to keep the focus on edits.
The x axis counts time in seconds from the start of the session.
Most sessions last a few dozen seconds~(\cref{t:session-size}), so the axis
ends at 150 seconds to show some but not all of the long tail.
The y axis shows changes to errors over time; a positive number means errors
increased since the last record, and a negative number represents a decrease.
While there are a few errors that go outside the plot bounds
(max \mnocheck{}: 51, max \mnonstrict{}: 3586, max \mstrict{}: 87), we focus
on $\pm{}50$ errors because most of the mass is in that region.
The points are color-coded by their session.
There are many colors througout the plot, which shows that the data comes
from many different sessions rather than from a few high-traffic sessions.

@@

The visual symmetry in the density plots~(\cref{f:error-density})
suggests that edits add and remove errors with rougly equal frequency,
regardless of mode.
\Cref{f:error-changes} explores this question of balance in detail.
For each of the three modes, it reports the percent of edits that increase (\addsym{}),
maintain (\keepsym{}), or decrease (\dropsym{}) the number of type
errors in the module.
It also reports counts for \FS{} checks to see whether fixing type errors fixes
\FS{} errors as well.
The ``maintain'' rows include cases where there are zero errors before and after.

@@

With the data in hand, we can answer the research questions
from~\cref{s:introduction}.
The answers confirm several aspects of Luau's design, but
at the same time offer several directions for future improvements.

@@

        Luau has an efficient type checker that can
        analyze thousands of lines of code~(\cref{t:codebase-size}) quickly
        enough to be used on-the-fly in an IDE.
        But, few sessions use type analysis directly.
        We conjecture that Luau needs further tailoring to match
        common idioms in untyped code.
        Discovering these idioms is a question for future work~(\cref{s:conclusion}).

@@

      \item[- How do sessions respond? Which errors persist through edits?]
        Sessions typically fix analysis errors.
        With few exceptions, increases and decreases in the number of errors
        balance out~(\cref{f:error-changes,t:type-error-survival,}).
        Even in \mnocheck{}, most edits that overlap with syntax errors
        remove at least one error~(\pct{67}) rather than increase or maintain
        the total.
        The errors that most often survive edits are due to
        inference at a binary operation,
        sending \code{self} to a function that does not expect it,
        and failure to unpack an optional value.
        But the first two seldom arise, and the third is something creators
        might reasonably ignore while prototyping.

>>>>

{Roblox} is a platform for {shared virtual experiences} (typically 3D games), with
70~million Daily Active Users, and 16~billion hours of engagement in
July--September 2023~\cite{corp.roblox.com}.
There are over 5~million distinct user-created programs available on the platform thanks to a
worldwide community of 3~million creators.
Many of the creators program for fun or learning and may
not consider themselves software developers; others are professional
developers who work full time with the Roblox platform.

@@

Due to the broad community of creators, the goals of the Luau type
system are rather unique~\cite{bfj-hatra-2021}.
Whereas a traditional type system focuses on compilation and memory
safety, Luau takes a lenient approach by default and lets creators
gradually~\cite{st-sfp-2006,tf-popl-2008} migrate to rigorous checks
one module at a time via three analysis modes.
The default \mnocheck{} mode reports only syntax errors,
the \mnonstrict{} mode reports would-be runtime errors and skips over
other issues,
and the \mstrict{} mode reports a variety of potential errors.
In addition, no matter the mode, a {\FS{}} type analysis runs to guide \IDE{} tools~(\cref{s:context}).
Luau's types aim to support untyped designs, in the spirit of migratory
typing~\cite{tfffgksst-snapl-2017}, so that creators can switch modes without
needing to restructure their code in major, potentially-breaking, ways.

@@

In this work, we design telemetry that collects data on type errors without
exposing source code, source locations, or even error message text (which
may contain revealing information).
The telemetry counts the number and \emph{kind} of type errors at various
levels of granularity.
Furthermore, it maintains a client-side approximation of the latest source-code
edits and uses that to identify type errors that overlap with this edit range.
Each record contains a \panon{} session identifier and a timestamp.

With this telemetry data, we investigate research questions about
the adoption and benefits of type analysis:
\begin{description}
  \item[RQ1.]
    How many sessions use type analysis?
    How often do projects contain modules with different
    analysis modes?
    How often do sessions turn analysis off?
  \item[RQ2.]
    For modules that use type analysis:
    which errors arise,
    how do sessions respond,
    and which errors tend to persist despite subsequent edits?
  \item[RQ3.]
    What impact does type analysis have on the number of \FS{}
    errors?
    For example, do \FS{} errors pile up in unanalyzed (\mnocheck{}) projects?
\end{description}

@@

At a higher level, this paper is the first to use telemetry 
to study a type checker.
It thus represents a step toward data-driven language design,
informed by many users' actual practice.
Our data captures over 340,000 sessions
that occured between February and April 2023
and covers thousands of type analysis errors
and millions of \FS{} errors.
By contrast to typical qualitative methods such as surveys and interviews, it
is not restricted to users' \emph{perceptions} about their work and it is not
limited to a small number of users performing synthetic tasks.

@@

  \item
    A dataset of over 1.5 million telemetry records and scripts to analyze them~\cite{luau-artifact}.

@@

Creators can opt in to detailed error reports and highlights by
choosing a \emph{type analysis mode} for each script.
The modes are:
\begin{itemize}
  \item \mnocheck{}: report only syntax errors (the default),
  \item \mnonstrict{}: report syntax errors and a subset of high confidence type errors, and
  \item \mstrict{}: report syntax errors and all type errors.
\end{itemize}
Each run of the type analysis can report several errors.
There is no guarantee that creators read every error.
In fact, creators who close the Script Analysis widget
can see highlights in their code but no further details about the errors.

@@

 As an example of \mnonstrict{} mode, the following program reports only one error:
\begin{verbatim}
  --!nonstrict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> OK
  local z = x.r           --> UnknownProperty: Key 'r' not found in table 'x'
\end{verbatim}
In \mstrict{} mode, it reports two errors:
\begin{verbatim}
  --!strict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> TypeMismatch: Type 'nil' could not be converted into 'number'
  local z = x.r           --> UnknownProperty: Key 'r' not found in table 'x'
\end{verbatim}
In programs like the above, where it is undecidable whether there will be a run-time error,
\mstrict{} mode errs on the side of reporting an error and \mnonstrict{} mode errs on
the side of suppressing the error.

Both modes report the \code{UnknownProperty} error because
misspellings of property names are common enough to report in both
\mstrict{} and \mnonstrict{} mode.
The \mnocheck{} mode reports no error because the program is syntactically valid.
See~\cite{bfj-hatra-2021} for a more detailed discussion of the rationale
behind the type systems.

In addition to the main type analysis, Roblox Studio takes a second pass over
every codebase with a \emph{\FS{}} analysis to infer autocomplete suggestions
and drive other \IDE{} tools.
To infer precise types,
the \FS{} analysis uses rigorous checks very similar to those of
\mstrict{} mode (but not identical: see \cref{s:strict-vs-forcedstrict})
no matter what mode the creator declared for the script.
When it fails to infer a type, \FS{} analysis defaults to the unknown gradual
type~\cite{st-sfp-2006}.
Creators cannot see the errors reported by \FS{} analysis.
Internally to Roblox Studio, this analysis is called \emph{forced strict}
because it ignores the creator's choice and applies strong checks.

\Cref{t:type-error-labels} lists a few of the type errors that
script analysis can report (10 out of 35 total)
to give a sense of Luau.
A \code{SyntaxError} is the only error that can appear in
\mnocheck{} code.
Several errors including \code{UnknownProperty} are about
tables, which are Luau's (and Lua's) primary data structure.
Tables encode arrays, dictionaries, and objects; consequently,
tables can have methods and can usually be extended.
The error \code{TypesAreUnrelated} is a special kind of type mismatch
that arises only during cast, unification, or subtyping check.
A \code{CountMismatch} occurs when the arguments to a function
do not match the function's arity.
Similarly, an \code{IncorrectGenericParamCount} is an arity mismatch
for a generic type.
A \code{CodeTooComplex} occurs when the typechecker hits an
internal limit on problem size. Lastly, \code{GenericError} is a catch-all
label for miscellaneous errors.
When Luau can provide more context regarding a generic error, it uses
a second error label called \code{ExtraInformation} to attach
a brief description.

@@

  \begin{tabular}{ll}
    Label & Interpretation \\\midrule
    \code{TypeMismatch} & Basic type error. \\
    \code{SyntaxError} & Basic parse error, e.g., \code{for if end}. \\
    \code{UnknownProperty} & Referenced an invalid field or method.  \\
    \code{OnlyTablesCanHaveMethods} & Tried to attach a method to a non-table. \\
    \code{CannotExtendTable} & Tried to extend a sealed table. \\
    \code{TypesAreUnrelated} & Failed cast, unify, or subtype. \\
    \code{CountMismatch} & Arity mismatch for a function. \\
    %% \code{CannotCallNonFunction} & Called a value that is not a function. \\
    \code{IncorrectGenericParamCount} & Arity mismatch for a generic type. \\
    \code{CodeTooComplex} & Type analysis failed~(\cref{s:code-too-complex}). \\
    %% \code{UnificationTooComplex} & Type analysis failed, unification solver hit a limit \\
    \code{GenericError}
    & Generic label for other non-type errors,\\
    & \hbox{}~~e.g., looping over an unordered table.
    %% attempting to extend a type that does not describe a class
    %% attempting to iterate over a table without a clear order

@@

Luau modules provide reusable libraries, which may be
{imported} by other scripts.
Imports therefore form a graph in general, though type analysis
removes edges to ensure an acyclic graph.
In \mstrict{} mode, it is an error to create an import cycle.

@@

While a program is under development, it is typical for the data
model to be edited.
% For example, instances may be added or renamed.
% --> Commented out to get a better page break.
Since the shape of the data model tree is reflected in
the type system, it is possible for these edits to introduce a huge number
of type errors across the project.

@@

Telemetry allows an application to phone home with data
summarizing usage patterns, such performance data, crash reporting,
and feature uptake. A typical use-case is in deciding whether an API can
be deprecated; without telemetry, it may be difficult to know how
popular the API is, but with telemetry it is straightforward.
Telemetry for programming languages can, however, be controversial.
See, for example, the lively discussion around telemetry in the Go
toolchain~\cite{golang-telemetry}.

@@

\end{enumerate}
See Cox~\cite{transparent-telemetry} for further discussion of
the privacy implications and tradeoffs of telemetry for programming languages.

@@

\Cref{f:tdata} shows the structure of telemetry records:
\begin{enumerate}
  \item
    \uid{}: a \panon{} identifier (15-digit random number) for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summary of type errors and \FS{} errors during the
    last two invocations of type analysis, and the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: list of specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

@@

\Cref{t:session-size} similarly outlines the size of sessions using two metrics:
real time in seconds between the first and last record, and the total number of
records.
The plots focus
on the lower-left fragment of the dataset, which again has a very long tail.
%% comment on long-tail record counts?
The longest sessions last several days and/or contain thousands of records,
while the shortest last a few milliseconds or include just one record.
%% 0 ms = 1 event
A median session runs on the order of minutes and consists of a few dozen events.

@@

The low fraction of errors in the edit range sends a mixed message:
\begin{itemize}
  \item
    On one hand, it may be that edits successfully remove errors.
    This is true, however, only for type errors and not for \FS{} errors.
    There are 11,479 more type errors from the previous analysis
    that overlap with the current edit range~(\pct{27} difference),
    but 172,678 fewer overlapping previous \FS{} errors~(\pct{15} difference).
    Since creators can see type errors but not \FS{}
    errors before making edits, we conclude that visibility
    makes a difference.
    \Cref{s:type-error-survival} explores edits and type errors in more detail.

  \item
    On the other hand, the low overlap between errors (whether current or previous)
    and edits says that edits rarely target highlighted code.
    There are several possible explanations:
    analysis might report several errors and the edits only target one or two of them;
    creators might ignore errors entirely and edit other code; or
    creators might choose to edit a non-highlighted location to fix an error.
    While non-privacy-protecting analyses have answered such
    questions in the past~\cite{mfk-sigcse-2011}, our data cannot narrow down an answer.

@@

%% mostly pure, rare mixes
Most sessions (\pct{99.82}) stick to a single analysis mode.
They never change the mode in the current module and never
switch focus to a module with a different analysis mode.
Among the other sessions, most of those (\pct{80}) switch to a
module with a different mode, about half contain at least one edit that
upgrades to a more strict mode, and about half contain a downgrade to a
less-strict mode.
There are 263 total upgrades across all sessions and 320 total downgrades.

@@

The data does not support this explanation.
On average, mode-upgrades resulted in 3 additional
type errors (stddev 7, median 0).
This data excludes module switch records and considers only keystroke events.
The worst-case increase was quite high, at 57 type errors, but exceptional.
In another exception, upgrading modes removed three type errors,
possibly due to other edits being rolled in with the
mode change.

@@

\Cref{f:error-by-mode} divides errors across modes.
The left plot shows that \mnonstrict{} accounts for two thirds of all type errors.
This confirms our expectation.
%% because \mnocheck{} reports only syntax errors and
%% \mstrict{} covers only a small number of telemetry records.
The right plot, by contrast, does not support our hypothesis that \mnocheck{}
records have more \FS{} errors than the other modes.
Though \mnocheck{} does have the most \FS{} errors, its \emph{proportion} matches
the proportion of \mnocheck{} records in the
dataset~(\cref{f:total-records})---if every record had exactly one \FS{} error,
the plot would be the same.

@@

The main focus of our telemetry design is to learn how type errors intersect
with edit ranges.
Indeed, most of the fields in a telemetry record are dedicated to this
topic~(\cref{f:tdata}), and are directed toward RQ2 on type errors.

@@

    \subitem Curiously, the number of syntax errors never decreases or stays the same.
    This may be because the easy fix is to press undo; telemetry does not
    see undo events.

    %% TODO awkward
    \subitem The number of typos is inflated by our telemetry strategy because it cannot
    determine when a creator is mid-edit.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will generate an
    \code{UnknownSymbol} error for \code{buck} and
    \code{UnknownProperty} for \code{bucket.c}, both of which may end up
    getting sampled.
    \Cref{s:conclusion} suggests ways
    to avoid reporting typos in the future.

  \item
    After the syntax errors, \code{CountMismatch}, \code{TypeMismatch}, and
    \code{GenericError} are the next most common \mnonstrict{} and \mstrict{}
    errors.
    A \code{CountMismatch} sends the wrong number of arguments to a function.
    It may simply be a common error, but it may also be a
    telemetry artifact.
    The high incidence of \code{TypeMismatch} and \code{GenericError} calls for
    further study in future work to determine precisely which error arose.
    Some of the \code{GenericError}s come with context strings in the form of
    \code{ExtraInformation} errors, but telemetry does not report these strings.
    %% GenericError = only a few things: extend non-class, iterate non-table, ...?

@@

  \item
    \code{TypeMismatch} is far more common in \mstrict{} mode (\pct{20.46})
    than \mnonstrict{} (\pct{2.44}) because \mstrict{}
    has additional type constraints.
    Several other errors have higher \mstrict{} percentages for the
    same reason, such as \code{CannotInferBinaryOperation}, and
    two errors appeared only in \mstrict{}:
    \code{OccursCheckFailed} and \code{IncorrectGenericParamCount}.

  \item
    Several Luau errors never appeared in edit ranges.
    Four of these deal with internal typechecker limits, which we discuss
    below.
    The others are for
    dynamic property lookups,
    deprecated APIs, swapped generic parameters,
    functions that require a self, and type pack mismatches.
    We would not expect these to appear often.

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{ktu-caap-1990,m-popl-1990}, the Luau typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors along with a message
that the analysis failed to understand the code:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
These errors can appear in Script Analysis just like any other
error, but one design goal of Luau is that code rarely hits the limits.
A typical creator should never see these errors.

@@

Since errors and edits may not overlap, either because the highlights
are misleading or because the creator ignores the errors, it is important
to see how the total number errors changes over time.
We measure \emph{error density} rather than error count, however,
to normalize for the wide range of codebase sizes~(\cref{t:codebase-size}).
Density is the number of type errors divided by the number of lines.

\Cref{f:error-density} plots changes to error density for every session in the dataset.
The data excludes module-switch telemetry records to keep the focus on edits.
The x axis counts time in seconds from the start of the session.
Most sessions last a few dozen seconds~(\cref{t:session-size}), so the axis
ends at 150 seconds to show a bit of the long tail.
The y axis shows changes to errors over time; a positive number means errors
increased since the last record, and a negative number represents a decrease.
While there are a few errors that go outside the plot bounds
(max \mnocheck{}: 51, max \mnonstrict{}: 3586, max \mstrict{}: 87), we focus
on $\pm{}50$ errors because most of the mass is in that region.
The points are color-coded by their session.
There are many colors througout the plot, which shows that the data comes
from many different sessions rather than from a few high-traffic sessions.

@@

The visual symmetry in the density plots~(\cref{f:error-density})
suggests that edits add and remove errors with rougly equal frequency,
regardless of mode.
\Cref{f:error-changes} explores this question of balance in detail.
For each of the three modes, it reports the percent of edits that increase (\addsym{}),
maintain (\keepsym{}), or decrease (\dropsym{}) the number of type
errors in the module.
It also reports counts for \FS{} checks to see whether fixing type errors fixes
\FS{} errors as well.
The ``maintain'' rows exclude cases where there are zero errors before and after.

@@

With the data in hand, we can answer the research questions
from~\cref{s:introduction}:

@@

        \subitem Luau has an efficient type checker that can
        analyze thousands of lines of code~(\cref{t:codebase-size}) quickly
        enough to be used on-the-fly in an IDE.
        But, few sessions use type analysis directly.
        We conjecture that Luau needs further tailoring to match
        common idioms in untyped code.
        Discovering these idioms is a question for future work~(\cref{s:conclusion}).

@@

      \item[- How do sessions respond? Which errors persist through edits?]
        Sessions appear to fix analysis errors.
        With few exceptions, increases and decreases in the number of errors
        balance out~(\cref{f:error-changes,t:type-error-survival,}).
        Even in \mnocheck{}, most edits that overlap with syntax errors
        remove at least one error~(\pct{67}) rather than increase or maintain
        the total.
        The errors that survive edits most often are due to
        inference at a binary operation,
        sending \code{self} to a function that does not expect it,
        and failure to unpack an optional value.
        But, the first two seldom arise and the third is something creators
        might reasonably ignore while prototyping.
